Quantum cluster approach to the topological invariants in correlated Chern insulators

### 2019-05-23 paper/paper.tex
Revised version of NJP.

\par We want to remark that the actual phase diagram of the Haldane Hubbard model in the moderate interacting regime has been highly debated over the past several years, namely different methods predicted quite different results \cite{HZKL_PRB2011-1,HZKL_PRB2011-2,MR_PRB2013,HRP_PRB2015,ZSWZ_PRB2015,WFSM_PRB2016,HCPP_PRL2016,VSLTHT_PRL2016,IWT_PRB2016,ASHP_PRB2016,GJMP_PRB2016,GR_NJP2018,LTTNNH_PBCM2018}. In particular, even whether the intermediate phases exist has not come to an agreement, let alone their physical properties. Due to the limited power of quantum cluster approaches, we have no intention to settle down this difficult problem in this paper. It is only safe to claim that in the framework of CPT/VCA, there exist an intermediate phase (NMI), which is not topological in the sense that it has no nonzero interacting bulk Chern number nor gapless states or zero-energy zeros of the single particle Green's function on edges. Nevertheless, this claim cannot rule out the possibility of the exotic nonmagnetic phases with topological fractionalized excitations \cite{HZKL_PRB2011-2,MR_PRB2013,HRP_PRB2015,HCPP_PRL2016} because it is not guaranteed that the nontrivial topology of the fractionalized modes must manifest itself in the single particle Green's functions. It is also possible that the NMI phase is just an artifact of the CPT/VCA method. Indeed, as can be seen in Fig. \ref{phase}(a), the NMI phases connects to the well known fictitious ``spin liquid'' phase in the simple honeycomb Hubbard model with $t'=0$ \cite{MLWAM_N2010,SOY_SR2012,AH_PRX2013,HD_PRL2013}, which arises from the nonequivalent treatment of the intracluster hoppings and the intercluster ones of the six-site tiling\cite{LRTR_PRB2014}. Yet the experiences of the case with $t'=0$ may not be definitely generalized to that with a finite $t'$. In view of the effective spin model incorporated with finite-$U$ effects, a nonzero $t'$ term introduces next nearest neighbor spin interactions and ring-exchange spin interactions, which could frustrate magnetic order and stabilize exotic quantum disordered phases \cite{HRP_PRB2015,HCPP_PRL2016}. Besides, charge fluctuations are known to be possible to suppress magnetic order and favor nonmagnetic phases in the moderate interacting regime \cite{SS_PRL2008,YKK_PRL2009,RLRT_PRL2015}. Further sophisticated numerical analyses, such as the large scale density matrix renormalization group simulations, are needed to determine the phases in the moderate interacting regime.

\par However, whether the NMI phase exists has nothing to do with the reported fault here caused by the standard periodization procedure that leads to a false interacting Chern number. In fact, as mentioned in Sec. \ref{MM}, the density of states and the grand potential are not affected by the approximation employed by the periodization procedure, thus, the single particle gap $\Delta$ and the antiferromagnetic moment $M$ are irrelevant to this approximation. Note that $\Delta$ and $M$ completely determines the phase diagram shown in Fig. \ref{phase}(a). Therefore, the cause that accounts for a false interacting topological Chern number is different from the one that may result in a possible fictitious NMI. Thus, the validity of the modified computing scheme of applying quantum cluster approaches to the calculation of interacting topological invariant is independent of the existence of this specific ``NMI'' phase. In fact, it applies to all insulating phases that could either be nonmagnetic or host magnetic orders. Indeed, based on the same computing scheme, the whole AFM phase is shown to be non-topological, which invalidates the claim in a previous quantum cluster study \cite{WFSM_PRB2016} based on the false natural-unit-cell Chern number that there exists a topological region in the AFM phase in the moderate interacting regime. Besides, the noninteracting spinless Haldane model has been realized in recent cold atom experiments using the shaking lattice technique \cite{JMDRLUGE_N2014}. We hope that our claim can be tested in future experiments.

>>>>

\par We want to remark that the actual phase diagram of the Haldane Hubbard model in the moderate interacting regime has been highly debated over the past several years, namely different methods predicted quite different results \cite{HZKL_PRB2011-1,HZKL_PRB2011-2,MR_PRB2013,HRP_PRB2015,ZSWZ_PRB2015,WFSM_PRB2016,HCPP_PRL2016,VSLTHT_PRL2016,IWT_PRB2016,ASHP_PRB2016,GJMP_PRB2016,GR_NJP2018,LTTNNH_PBCM2018}. In particular, even whether the intermediate phases exist has not come to an agreement, let alone their physical properties. Due to the limited power of quantum cluster approaches, we have no intention to settle down this difficult problem in this paper. It is only safe to claim that in the framework of CPT/VCA, there exist an intermediate phase (NMI), which is not topological in the sense that it has no nonzero interacting bulk Chern number nor gapless states or zero-energy zeros of the single particle Green's function on edges. Yet, this claim cannot rule out the possibility of the exotic nonmagnetic phases with topological fractionalized excitations \cite{HZKL_PRB2011-2,MR_PRB2013,HRP_PRB2015,HCPP_PRL2016} because it is not guaranteed that the nontrivial topology of the fractionalized modes must manifest itself in the single particle Green's functions. It is also possible that the NMI phase is just an artifact of the CPT/VCA method. Indeed, as can be seen in Fig. \ref{phase}(a), the NMI phase connects to the well known fictitious ``spin liquid'' phase in the simple honeycomb Hubbard model with $t'=0$ \cite{MLWAM_N2010,SOY_SR2012,AH_PRX2013,HD_PRL2013}, which arises from the nonequivalent treatment of the intracluster hoppings and the intercluster ones of the six-site tiling\cite{LRTR_PRB2014}. Yet the experiences of the case with $t'=0$ might not be definitely generalized to that with a finite $t'$. In view of the effective spin model obtained in the large-$U$ expansion, a nonzero $t'$ term introduces next nearest neighbor spin interactions and ring-exchange spin interactions, which could frustrate magnetic order and stabilize exotic quantum disordered phases \cite{HRP_PRB2015,HCPP_PRL2016}. Besides, charge fluctuations are known to be possible to suppress magnetic order and favor nonmagnetic phases in the moderate interacting regime \cite{SS_PRL2008,YKK_PRL2009,RLRT_PRL2015}. Further sophisticated numerical analyses, such as the large scale density matrix renormalization group simulations, are needed to determine the phases in the moderate interacting regime. We note that the noninteracting spinless Haldane model has been realized in recent cold atom experiments using the shaking lattice technique \cite{JMDRLUGE_N2014}. We hope that our claim can be tested in future experiments.

\par We emphasize, whether the NMI phase exists is in fact not directly related to the reported fault here caused by the standard periodization procedure that leads to a false interacting Chern number. As mentioned in Sec. \ref{MM}, the density of states and the grand potential are not affected by the approximation employed by the periodization procedure, thus, the single particle gap $\Delta$ and the antiferromagnetic moment $M$ are irrelevant to this approximation. Note that $\Delta$ and $M$ completely determine the phase diagram shown in Fig. \ref{phase}(a). Therefore, the cause that accounts for a false interacting topological Chern number is different from the one that may result in a possible fictitious NMI. Thus, the validity of the modified computing scheme of applying quantum cluster approaches to the calculation of interacting topological invariant is independent of the existence of the ``NMI'' phase. In fact, it applies to all insulating phases that could either be nonmagnetic or host magnetic orders. Indeed, based on the same computing scheme, the whole AFM phase is shown to be non-topological, which invalidates the claim in a previous quantum cluster study based on the false natural-unit-cell Chern number that there exists a topological region in the AFM phase in the moderate interacting regime.
